Boutiques: a flexible framework for automated application integration in computing platforms

### 2016-10-19 paper.tex
More text

\documentclass{article}
\usepackage{minted}

\newcommand{\note}[2]{\pdfmargincomment[color=yellow,author=#1,open=true]{#2}}

@@

Why Linux. Define application. Define computing platforms. Explain the problem and why it is difficult.

@@

\paragraph{General properties.} Inputs must have a name, a unique
identifier and a type. They may be optional, have a description, a
command-line key (placeholder), a flag and flag separator, and a
default value. Inputs may also be lists.

\paragraph{Types.} Inputs may be of type \texttt{String},
\texttt{Number}, \texttt{Flag} or \texttt{File}. Types can be
restricted to a specific set of values (\texttt{String} and
\texttt{Number} only), to a specific range (\texttt{Number} only) to
integers or booleans (\texttt{Number} only).

@@

@@

@@

@@

@@

@@

% cbrain-plugins-neuro
% PSOM
% MICCAI Challenges
% exports from Nipype
% SPM?

>>>>

\documentclass{article}
\usepackage{minted}
\usepackage{pdfcomment}

\newcommand{\note}[2]{\pdfmargincomment[color=yellow,author=#1,open=true]{#2}}
\newcommand{\todo}[2]{\pdfmargincomment[color=red,author=#1,open=true]{#2}}

@@

Why Linux. Define application. Define computing platforms. Explain the
problem and why it is difficult. Approach: start simple and enrich the
spec when new applications come. Refer to computing infrastructures
and the need for parallelization. Define task vs. invocation. Who are
the developers. Refer to neuroinformatics field.

@@

\paragraph{General properties.} Inputs must have a name, a unique
identifier and a type. They may be optional, have a description, a
command-line key (placeholder), a flag and flag separator, and a
default value. Inputs may also be ordered lists.

\paragraph{Types.} Inputs may be of type \texttt{String},
\texttt{Number}, \texttt{Flag} or \texttt{File}. \texttt{File} may
also represent a directory. Types can be restricted to a specific set
of values (\texttt{String} and \texttt{Number} only), to a specific
range (\texttt{Number} only) to integers or booleans (\texttt{Number}
only).

@@

Application outputs are the files or directories that need to be
transferred and delivered to the user once the execution is
complete. While it may not be required to distinguish inputs and
outputs to build the application command line, this information is
needed by computing platforms to determine what needs to be saved from
the execution.

In Boutiques, output files are defined from a unique identifier, a
name and a path template that specifies the file or directory
name. Path templates may include input command-line keys when output
files are named after the input values. Input values may be stripped
from specific strings before being substituted in the path template,
e.g., file extensions. Output files may also have a description, a
command-line flag, a flag separator, and a command-line key in case
they appear on the command line. They may be optional in case the file
is not always produced by the application (for instance when it is
produced only when a particular flag is activated). They may also be
unordered lists; in this case, the path template must contain a
wildcard (\texttt{*}) standing for any string of characters (as in
\texttt{bash}) and defining the pattern used to match the output files
in the list.

Listing~\ref{listing:output-example} shows the definition of an output
file in the command line exemplified before. According to this
definition and assuming that the string input value entered by the
user is "foo.csv", the string \texttt{[LOG]} will be
replaced by \texttt{log-foo} on the command line at execution time.

\begin{listing}
\begin{minted}[frame=single,
               framesep=3mm,
               linenos=false,
               xleftmargin=21pt,
               tabsize=4]{js}

{
    "id" : "logfile",
    "name" : "Log file",
    "description" : "The output log file from the example tool",
    "path-template" : "log-[STRING_INPUT]",
    "command-line-key" : "[LOG]",
    "path-template-stripped-extensions" : [".txt", ".csv"],
    "optional" : false
}
\end{minted}
\caption{Output file example} 
\label{listing:output-example}
\end{listing}

\subsection{Invocation schema}
\label{sec:invocation-schema}

Input validation is an important motivation for common application
descriptions. The obvious solution in our context is to rely on an
application-specific JSON schema, called ``invocation schema'' to
specify the correct input values that an application
accepts. Platforms could leverage the invocation schema to validate
inputs automatically and without having to develop specific code,
using any JSON validator.

Invocation schemas, however, are complex JSON objects. Basically, they
must represent the properties described above in a formal way,
including dependencies between inputs. To illustrate such complexity,
listing~\ref{listing:invocation-schema-example} displays
\todo{Tristan}{...}. % must support dependent invocation ids to allow for workflow support.

It does not seem realistic to assume that
developers will easily write correct JSON schemas for their
applications. In addition, invocation schemas can actually be
generated automatically from the basic application description. Thus,
the invocation schema is an optional property of the Boutiques schema,
that platforms or developers may generate automatically using a core
tool provided by Boutiques.

@@

A large number of applications actually rely on configuration files
rather than command-line options to define their input and output
parameters. Indeed, when the number of parameters increases,
command-line may become long and cumbersome while configuration files
allow for better structure and documentation.

Configuration files may be complex though, and specified in any
language.  As it does not seem reasonable to specify a particular
template for configuration files, the Boutiques specification allows
application developers to specify their own template, referring to
input and output command-line keys \todo{Tristan}{Then they are not
  command-line keys any more} to specify placeholders for input and
output values. Listing~\ref{listing:configuration-file-example} shows
an example. Configuration files also have a path template that defines
how they must be named and where they must be written. They may also
have a command line key and flag in case they need to be passed on the
command-line. 
\todo{Tristan}{See Github issue \#17}

@@

Boutiques intentionally does not support the definition of complete
workflows from individual applications. Indeed, we believe that
workflow engines are specific tools that need complex languages to
implement the particular functions that they target. Efforts to
specify common workflow languages include the Common Workflow Language
(CWL)~\cite{cwl} and the Interoperable Workflow Intermediate
Representation (IWIR)~\cite{plankensteiner2011iwir}.

Instead, Boutiques allows to wrap complete workflows as simple
applications depending on particular workflow engine(s) installed in
container images. In this model, workflows are executed as any other
task on the infrastructure, which allows for basic data parallelism by
iterating the workflow on a set of independent files (for instance:
subjects in an analysis). Data parallelism, however, may not be
sufficient when the analysis of a particular file has to be
parallelized. To address this issue, a specific property of the
Boutiques schema specifies that an application may submit new
tasks. New tasks are created by the application by writing a JSON
object in the execution directory, complying to the invocation schema
described in Section~\ref{sec:invocation-schema}. The platform must
read this object, submit the corresponding task, and write back the
invocation identifier in the execution directory. The invocation
identifier may be used by the workflow to specify dependencies between
tasks in subsequent task submissions.

The adopted ``workflow encapsulation'' model is very simple but it
allows for a scalable and reliable execution of workflows expressed in
a variety of languages, as detailed in ~\cite{glatard2016fgcs}.

@@

Containers are adopted because they allow for an automated and
lightweight integration of application implementations in
platforms. They are extremely useful to improve the reproducibility of
analyses as variations in the software environment may have an
important impact on the computed results. They also have limitations,
in particular they do not specify the hardware architecture required
to execute an application, which can be an issue in some cases.

@@

Custom properties may be added to the Boutiques specification without
restriction. They may be useful to implement platform-specific
features but they should be used with care to avoid making tools
dependent on a particular platform. The following custom properties
have been used so far:
\begin{itemize}
\item \texttt{cbrain:can-submit-new-tasks}: a boolean, true if the tool may submit new tasks to the platform. \todo{Tristan}{related to the workflow support section}
\item \texttt{cbrain:inherits-from-class}: a string that defines the Ruby class that should be used as parent class for the
 tool in CBRAIN. Used to define a progress bar for PSOM tools in CBRAIN.
\item \texttt{vip:miccai-challenger-email} and
  \texttt{vip:miccai-challenge-team-name}: strings helping VIP
  categorize tools for the 2016 MICCAI challenges ``MSSEG'' and
  ``PETSEG''.
\end{itemize}

\subsection{Command-line construction}

@@

% cbrain-plugins-neuro
% PSOM
% MICCAI Challenges
% exports from Nipype
% SPM?
% BIDS Apps
